---
title: "Movie Budgets and Revenues"
author: ""
date: "08.17.20"
output: 
  html_document:
    toc: true
    toc_float: true
---

We will look at the relationship between budget and revenue for movies made in the United States in 1986 to 2020. The dataset is created based on data from the [Internet Movie Database (IMDB)](https://www.imdb.com/).


```{r include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r load-packages, warning=FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(viridis)
library(DT)
```

## Data

The `movies` data set includes basic information about each movie including budget, genre, movie studio, director, etc. A full list of the variables may be found [here](https://github.com/danielgrijalva/movie-stats).

```{r read-data, message=FALSE, warning=FALSE}
movies <- read_csv("https://raw.githubusercontent.com/danielgrijalva/movie-stats/master/movies.csv")
```

View the first 10 rows of data. 

```{r}
movies %>%
  slice(1:10)
```


## Analysis

### Part 1 

We begin by looking at how the average gross revenue (`gross`) has changed over time. Since we want to visualize the results, we will choose a few genres of interest for the analysis. 

```{r genre-list}
genre_list <- c("Comedy", "Action", "Animation", "Horror")
```

```{r plot-rating, fig.width=10, warning=FALSE}
movies %>%
  filter(genre %in% genre_list) %>% 
  group_by(genre,year) %>%
  summarise(avg_gross = mean(gross)) %>%
  ggplot(mapping = aes(x = year, y = avg_gross, color=genre)) +
    geom_point() + 
    geom_line() +
    ylab("Average Gross Revenue (in US Dollars)") +
    ggtitle("Gross Revenue Over Time") +
    scale_color_viridis_d()
```

1. What do you observe from the plot? Type your responses in Ed Discussion. 


### Part 2

Next, let's see the relationship between a movie's budget and its gross revenue.

```{r plot-revenue, fig.width=10, warning=FALSE}
movies %>%
  filter(genre %in% genre_list, budget > 0) %>% 
  ggplot(mapping = aes(x=log(budget), y = log(gross), color=genre)) +
  geom_point() +
  geom_smooth(method="lm",se=FALSE) + 
  xlab("Log-transformed Budget")+
  ylab("Log-transformed Gross Revenue") +
  facet_wrap(~ genre) + 
  scale_color_viridis_d()
```


1. Suppose we fit a regression model for each genre that uses budget to predict gross revenue. What are the signs of the correlation between `budget` and `gross` and the slope in each regression equation?

2. Suppose we fit the regression model from the previous question. Which genre would you expect to have the smallest residuals, on average (*residual = observed revenue - predicted revenue*)? 

3. In the remaining time, discuss the following: Notice in the graph above that `budget` and `gross` are log-transformed. Why are the log-transformed values of the variables displayed rather than the original values (in U.S. dollars)? 

## References

1. [https://github.com/danielgrijalva/movie-stats](https://github.com/danielgrijalva/movie-stats)
2. [Internet Movie Database](https://www.imdb.com/)

## Appendix

Below is a list of genres in the data set:

```{r}
movies %>% 
  arrange(genre) %>% 
  select(genre) %>%
  distinct() %>%
  datatable()
```
