---
title: "Tutorial 02 - Linear Regression"
author: "Olivia Fan"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    css: "./tutorials.css"
    toc: true
    toc_float: true
---

<!-- Setup -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE,
                      message=FALSE)
```

## Topics 

- Use simple linear regression to describe the relationship between a quantitative predictor and quantitative response variable.

- Estimate the slope and intercept of the regression line using the least squares method.

- Interpret the slope and intercept of the regression line.

## 1. Libraries

The ``library()`` function is used to access functionality that is provided by R packages, but is not included in base R.  ``install.packages()`` can be used to install new packages. Run this command from the console.  

```{r}
# install.packages("ggplot2")
```

First, load the package ``ggplot2`` that will be used throughout the tutorial for data visualizations.

```{r}
library(ggplot2)
```


## 2. Exploratory Data Analysis

This tutorial will be using the ``nhanes`` dataset where the variables are described in the file `nhanes-codebook.txt`. Load this data with the `load` function and specify the rda data file. 

```{r}
load(file='nhanes1518.rda')
```

The functions ``head()`` and ``names()`` can be used to explore the data.

```{r}
head(nhanes1518)
names(nhanes1518)
```

We will first explore the dataset by doing exploratory data analysts on the predictor variable `BMXWAIST`:

```{r}
# Basic histogram
ggplot(nhanes1518, aes(x=BMXWAIST)) + geom_histogram()+labs(x = "waist circumference", title = "distribution of waist circumference")
# Change the width of bins
ggplot(nhanes1518, aes(x=BMXWAIST)) + 
  geom_histogram(binwidth=1)+labs(x = "waist circumference", title = "distribution of waist circumference")
# Change colors
p<-ggplot(nhanes1518, aes(x=BMXWAIST)) + 
  geom_histogram(color="black", fill="white")+labs(x = "waist circumference", title = "distribution of waist circumference")
```

We see that the distribution of waist circumference is bimodal, with peaks at waist circumference of around 50 and 100 inches respectively.

We can also add the mean line, as well as overlay with transparent density plot. The value of alpha controls the level of transparency:

```{r}
# Add mean line
p+geom_vline(aes(xintercept=mean(BMXWAIST)),
            color="blue", linetype="dashed", size=1)
# Histogram with density plot
ggplot(nhanes1518, aes(x=BMXWAIST)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") 
```


## 3. Simple Linear Regression

We'll start with a fitting a simple linear model using the ``lm()`` function.  Instead of attaching the ``nhanes1518`` dataset, we also can specify the data from the ``lm()`` function. In the ``lm()`` function, the first variable is the response variable and the variables to the right of the ``~`` symbol are the predictor variable(s).

```{r}
lm.fit <- lm(BMXBMI ~ BMXWAIST, data = nhanes1518)
```


There are several ways that we can examine the model results.  First, we can just call the name of the ``lm()`` model for a brief summary.

```{r}
lm.fit
```

We can also use the ``names()`` function to list all of the names of variables in the ``lm.fit`` model:

```{r}
names(lm.fit)
```

The ``summary()`` function gives a more extensive overview of the model fit:

```{r}
summary(lm.fit)
```

The coefficients of the linear regression model can be extracted using the ``coef()`` function and the confidence interval(s) with the ``confint()`` function.

```{r}
coef(lm.fit)
confint(lm.fit)
```

We can use the ``predict()`` function to obtain prediction intervals or confidence intervals for a given value of the predictor variable, ``BMXWAIST``.  Note that when using the predict function, the column names and format of the new points at which to predict needs to be the same as the original data frame used to fit the ``lm()`` model.  If you encounter errors using the ``predict()`` function, this is a good first thing to check.

```{r}
predict(lm.fit, data.frame(BMXWAIST = (c(70, 80, 90))), interval = "confidence")
predict(lm.fit, data.frame(BMXWAIST = (c(70, 80, 90))), interval = "prediction")
```


We can plot the variables ``BMXBMI`` and ``BMXWAIST`` using the ``plot()`` function and overlay the regression line found using ``lm()`` with the ``abline()`` function.

```{r}
plot(x=nhanes1518$BMXWAIST,y=nhanes1518$BMXBMI, xlab="Body Mass Index (kg/m**2)",ylab="Waist Circumference (cm)")
abline(lm.fit)
```

Here the dollar sign specifies that we retrieve the predictor columns `BMXBMI` and `BMXWAIST` from the dataset `nhanes1518`

We can experiment with different options for ``abline()`` by changing the line width and color in ``abline()``.

```{r}
plot(x=nhanes1518$BMXWAIST,y=nhanes1518$BMXBMI, xlab="Body Mass Index (kg/m**2)",ylab="Waist Circumference (cm)")
abline(lm.fit, lwd = 3, col = "red")
```

```{r}
plot(x=nhanes1518$BMXWAIST,y=nhanes1518$BMXBMI, xlab="Body Mass Index (kg/m**2)",ylab="Waist Circumference (cm)", col = "red")
```

The ``pch`` argument in ``plot()`` changes the shape/type of the points that are plotted.

```{r}
plot(x=nhanes1518$BMXWAIST,y=nhanes1518$BMXBMI, xlab="Body Mass Index (kg/m**2)",ylab="Waist Circumference (cm)",pch = 20)
```

```{r}
plot(x=nhanes1518$BMXWAIST,y=nhanes1518$BMXBMI, xlab="Body Mass Index (kg/m**2)",ylab="Waist Circumference (cm)",pch = "+")
```


Optional: We can make a similar plot using ``ggplot``, where we fit the linear regression model using ``ggplot()``.

```{r}
library(ggplot2)
ggplot(nhanes1518, aes(y = BMXBMI, x = BMXWAIST)) + 
    geom_smooth(method = "lm", formula = y ~ x, colour = "blue") + 
    geom_point() +
  ggtitle("Waist circumference vs. BMI for the nhanes data")

```

We can use the ``residuals()`` and ``rstudent()`` functions to extract the residuals and studentized residuals, respectively, from the linear model and plot them along with the predicted values.

```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

Additionally, we can compute the influence matrix for the predictors using the ``hatvalues()`` function.

```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

## 4. Model Diagonostics & Interpretation
The ``par()`` function can be used to create a grid of multiple subplots.

```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

The diagnostic plots show residuals in four different ways:

- Residuals vs Fitted. Used to check the linear relationship assumptions. A horizontal line, without distinct patterns is an indication for a linear relationship, what is good. The model we fitted shows roughly a linear relationship, with no distinct patterns (such as a fan or funnel shape) in the residuals vs. fitted plot. 

- Normal Q-Q. Used to examine whether the residuals are normally distributed. It’s good if residuals points follow the straight dashed line. The Q-Q plot generally follows the straight dashed line, with some deviations at the end towards high values of theoretical quantiles.

- Scale-Location (or Spread-Location). Used to check the homogeneity of variance of the residuals (homoscedasticity). Horizontal line with equally spread points is a good indication of homoscedasticity. 

- Residuals vs Leverage. Used to identify influential cases, that is extreme values that might influence the regression results when included or excluded from the analysis. Based on the residuals vs. leverage plot, there are no influential points according to Cook's distance. However, there might be some points with high standard residuals values which could be marked as outliers.

Some other metrics from the model:

- $R^2$: From the model above, we have an adjusted R-squared value of 0.8744, which indicates that 87.44% of the variability in the response variable Body BMI can be explained by the change in the predictor variable waist circumference. 
- p value: The p value tells us how likely the data we have observed is to have occurred under the null hypothesis (more material on Null hypothesis on subsequent tutorials), i.e. that there is no correlation between the predictor variable waist circumference and the response BMI. From the model above, we have a p value of 2.2e-16, which tells us that the predictor variable waist circumference is statistically significant. 

## 5. Multiple Linear Regression

The ``lm()`` function can also fit multiple regression models. In this section, we will use ``BMXWAIST``, `BMXWT`` and ``BMXHT`` as predictors of the response variable ``BMXBMI``.


```{r}
lm.fit <- lm(BMXBMI ~ BMXWAIST + BMXWT + BMXHT, data = nhanes1518)
summary(lm.fit)
```

In the lm() formula, a dot . can be used to include all variables in the Boston data as predictors.
```{r}
library(dplyr)
nhanes_na_removed<-cbind(nhanes1518[1:5],nhanes1518$BMXBMI)
# data clearning to ignore NA values
lm.fit1 <- lm(nhanes1518$BMXBMI ~ ., data = nhanes_na_removed)
summary(lm.fit1)
```

If we want to exclude specific variables from the list of predictors, we can use the ``-`` notation.  In the following example, all predictor variables but ``age`` are included in the model.

```{r}
library(dplyr)
nhanes_na_removed<-cbind(nhanes1518[1:5],nhanes1518$BMXBMI)
# data clearning to ignore NA values
lm.fit1 <- lm(nhanes1518$BMXBMI ~ . - RIDSTATR, data = nhanes_na_removed)
summary(lm.fit1)
```

Including ``-1`` excludes the intercept from the model.  
```{r}
lm.fit1 <- lm(nhanes1518$BMXBMI ~ .- 1, data = nhanes_na_removed)
summary(lm.fit1)
```

The ``update()`` function can be used to specify a new formula for an existing model.

```{r}
lm.fit1 <- update(lm.fit, ~. - RIDSTATR)
```


## 6. Interaction Terms

There are two ways to include interaction terms in the model, ``:`` and ``*``. The ``:`` symbol only includes the interaction term between the two variables, while the ``*`` symbol includes the variables themselves, as well as the interaction terms. This means that ``BMXWT*BMXWAIST`` is equivalent to ``BMXWT + BMXWAIST + BMXWT:BMXWAIST``.

```{r}
summary(lm(BMXBMI ~  BMXWT* BMXWAIST, data = nhanes1518))
```


A simple way to include all interaction terms is the syntax ``.^2``.
```{r}
library(dplyr)
nhanes_na_removed<-cbind(nhanes1518[1:5],nhanes1518$BMXBMI)
# data clearning to ignore NA values
summary(lm(nhanes1518$BMXBMI ~ .^2, data = nhanes_na_removed))
```

## Related Topics

| **Topics** | **Concepts** |
|-------------------------------------------------------------|:--:|
| ANOVA | Null hypothesis |
| Quantile factoring | use qualitative  as levels |
| Cross-validation 	|  k-fold, LOOCV	|
| Model section 	|  AIC, BIC, Cp	|

